\Section{Analysis}

\Paragraph{Reliability and Performance}

The three improvements described above resulted in a major qualitative
change in performance and reliability. In the version of \MVP released
in September 2020, correct examples verified fairly quickly and
reliably.  But that is because we needed speed and reliability, so we
disabled some properties that always timed out and others that timed
out unpredictably when there were small changes in the framework.  We
learned that incorrect programs or specifications would time out
predictably enough that it was a good bet that examples that timed out
were erroneous.  However, localizing the error to fix it was
\emph{very} hard, because debugging is based on a counterexample that
violates the property, and getting a counterexample requires
termination!

With each of the transformations described, we witnessed significant speedups
and, more importantly, reductions in timeouts.  Monomorphization was the last
feature implemented, and, with it, timeouts almost disappeared. Although this
was the most important improvement in practice, it is difficult to quantify
because there have been many changes in Diem framework, its specifications,
\MVP, and even the Move language over that time.

It is simpler (but less important) to quantify the changes in run time of \MVP
on one of our more challenging modules, the |DiemAccount| module, which is the
biggest module in the Diem framework. This module implements basic functionality
to create and maintain multiple types of accounts on the blockchain, as well as
manage their coin balances. It was called |LibraAccount| in release 1.0 of
\MVP. The table below lists the number of lines, functions, invariants,
conditions (requires, ensures, and aborts-if), as well as the verification
times:

{
\setlength{\tabcolsep}{6pt}
\vspace{2ex}
\begin{tabular*}{0.9\textwidth}{cccccc}
  \hline
  \hline
  Module & Lines & Functions & Invariants & Conditions & Timing \\
  \hline
  LibraAccount & 1975 & 72 & 10 & 113 & \textbf{9.131s} \\
  DiemAccount & 2554 & 64 & 32 & 171 & \textbf{6.290s} \\
  \hline
\end{tabular*}
\vspace{2ex}
}

\noindent Notice that |DiemAccount| has significantly grown in size compared to
the older version. Specifically, additional specifications have been
added. Moreover, in the original |LibraAccount|, some of the most complex
functions had to be disabled for verification because the old version of \MVP
would time out on them. In contrast, in |DiemAccount| and with the new version,
all functions are verified. Verification time has been improved by roughly 30\%,
in the presence of three times more global invariants, and 50\% more function
conditions.

We were able to observe similar improvements for the remaining of the 40 modules
of the Diem framework. All of the roughly half-dozen timeouts in verification in
the framework resolved after introduction of the transformations described in
this paper. Also, specifications which were introduced after the new
transformations did not introduce new timeouts.

%% A further improvement is that
%% often, in cases where specification and program disagree, a timeout occurred
%% which went away only after fixing the spec or the code, making debugging of such
%% verification failures rather hard. This problem also disappeared.


\Paragraph{Causes for the Improvements}

It's difficult to pin down and measure exactly why the three transformations
described improved performance and reliability so dramatically.  We have
explained some reasons in the subsections above: the alias-free memory model
reduced search through combinatorial sharing arrangments, and the fine-grained
invariant checking results in simpler formulas for the SMT solver.
%We believe that one reason is just that the translated SMT formulas are simpler.

We found that most timeouts in specifications stemmed from our liberal use of
quantifiers.  To disprove a property $P_0$ after assuming a list of properties,
$P_1, \ldots p_n$, the SMT solver must show that
$\neg P_0 \wedge P_1 \wedge \ldots \wedge P_n$ is satisfiable.  The search
usually involves instantiating universal quantifiers in $P_1, \ldots, P_n$.  The
SMT solver can do this endlessly, resulting in a timeout. Indeed, we often found
that proving a post-condition false would time out, because the SMT solver was
instantiating quantifiers to find a satisfying assignment of
$P_1 \wedge \ldots \wedge P_n$.  Simpler formulas result in fewer intermediate
terms during solving, resulting in fewer opportunities to instantiate quantified
formulas.

We believe that one of the biggest impacts, specifically regards
removal of timeouts and predictability of verification, is monomorphization. The
reason for this is that monomorphization allows a multi-sorted representation
of values in Boogie (and eventually the SMT solver). In contrast, before
monomorphization, we used a universal domain for values in order to represent
values in generic functions, roughly as follows:

\begin{Move}
  type Value = Num(int) | Address(int) | Struct(Vector<Value>) | ...
\end{Move}

\noindent This creates a large overhead for the SMT solver, as we need to
exhaustively inject type assumptions (e.g. that a |Value| is actually an
|Address|), and pack/unpack values. Consider a quantifier like~%
|forall a: address: P(x)| in Move. Before monomorphization, we have to represent
this in Boogie as~%
|forall a: Value: is#Address(a) => P(v#Address(a))|. This quantifier is
triggered where ever |is#Address(a)| is present, independent of the structure of
|P|. Over-triggering or inadequate triggering of quantifiers is one of the
suspected sources of timeouts, as also discussed in~\cite{BUTTERFLY}.

Moreover, before monomorphization, global memory was indexed in Boogie by an
address and a type instantiation. That is, for |struct R<T>| we would
have one Boogie array |[Type, int]Value|. With monomorphization, the type index
is eliminated, as we create different memory variables for each type
instantiation.  Quantification over memory content works now on a one-dimensional
instead of an n-dimensional Boogie array.

\Paragraph{Discussion and Related Work}

Many approaches have been applied to the verification of smart contracts
\cite{liu2019survey,miller2018smart,solcverify,DBLP:conf/esop/HajduJ20,verisol}.
A recent survey~\cite{CONTRACT_VERIFICATION} distinguishes between
\emph{contract} and \emph{program} level approaches. Our approach has aspects of
both: we address program level properties via pre/post conditions, and contract
(``blockchain state'') level properties via global invariants. In both cases, we
use traditional predicate logic to write these properties, characterized as
Hoare logic by the paper.

While~\cite{CONTRACT_VERIFICATION} refers to at least two dozen systems for
smart contract verification, to the best of our knowledge, the Move ecosystem is
the first one where contract programming and specification language are fully
integrated, and the language is designed from first principles influenced by
verification. Methodologically, Move and the Move prover are therefore closer to
systems like Dafny~\cite{DAFNY}, or the older Spec\# system~\cite{SPECSHARP},
where instead of adding a specification approach posterior to an existing
language, it is part from the beginning. This allows us not only to deliver a
more consistent user experience, but also to make verification technically
easier by curating the programming language, as reflected in Move's absence of
dynamic dispatching and the notorious re-entrance problem~\cite{REENTRANCE}, as
well as the borrow semantics which enables optimizations like reference
elimination (Sec.~\ref{sec:RefElim}).

% Other related works on the
% \solidity smart contract verification employ the theoretical foundations
% including the \kay framework \cite{kay} (e.g., \cite{kevm}), \fstar
% \cite{dm4all} (e.g., \cite{DBLP:conf/ccs/BhargavanDFGGKK16,post2018}), and proof
% assistants such as Coq \cite{coqref} (e.g., \cite{fspvme,fether}).

In contrast to other approaches that only focus on specific vulnerability
patterns~\cite{mythril,oyente,maian,securify}, \MVP offers a universal
specification language. To the best of our knowledge, no existing specification
approach for smart contracts based on inductive Hoare logic has similar
expressiveness. We support universal quantification over arbitrary memory
content, a suspension mechanism of invariants to allow non-atomic construction
of memory content, and generic invariants.  For comparison, the SMT Checker
build into \solidity \cite{solidity} does not support quantifiers, because it
interprets programming language constructs (requires and assert statements) as
specifications and has no dedicated specification language. While in Solidity
one can simulate aspects of global invariants using modifiers by attaching
pre/post conditions, this is not the same as our invariants, which are
guaranteed to hold independent of whether a user may or (accidentally) may not
attach a modifier, and which are optimized to be only evaluated as needed.

While the expressiveness of Move specifications and \MVP comes with the price of
undecidability and the dependency from heuristics in SMT solvers, \MVP is
capable to deal with this by its elaborated translation to SMT logic, as
partially described in this paper. The result is a practical verification system
that is fully integrated into the Diem blockchain production process, which (to
the best of our knowledge) is novel by itself.

Because of the similarity of Move and Rust~\cite{rust} in the use of borrow
semantics, some of the techniques described are also applicable to the
verification of Rust (e.g.~\cite{prusti,smack,nopanic,crust}).  For example, we
expect reference elimination to work similar for the safe subset of Rust.  To
the best of our knowledge, this has not been attempted before.

\Paragraph{Future Work}

%% DD Note: I'm taking a less specific and longer-term view of future work.
%% This is all a matter of taste, so feel free to go back to the previous
%% version if you prefer.

\MVP is conceived as a tool for achieving higher assurance systems, not as a bug
hunting tool. Having at least temporarily achieved satsifactory performance and
reliability, we are turning our attention to the question of the goal of higher
assurance, which raises several issues.  If we're striving for high assurance,
it would be great to be able to measure progress towards that goal.  Since
system requirements often stem from external business and regulatory needs,
lightweight processes for exposing those requirements so we know what needs to
be formally specified would be highly desirable.

As with many other systems, it is too hard to write high-quality specifications.
Our current specifications are more verbose than they need to be, and we are
working to require less detailed specifications, especially for individual
functions.  We could expand the usefulness of \MVP for programmers if we could
make it possible for them to derive value from simple reusable specifications.
Finally, software tools for assessing the consistency and completeness of formal
specifications would reduce the risk of missing bugs because of specification
errors.

However, as more complex smart contracts are written and as more people write
specifications, we expect that the inherent computational difficulty of solving
logic problems will reappear, and there will be more opportunities for improving
performance and reliability.  In addition to translation techniques, it will be
necessary to identify opportunities to improve SMT solvers for the particular
kinds of problems we generate.

%% While currently \MVP is performing to our satisfaction, there are multiple open
%% problems as well. For one, the set of smart contracts we are verifying still
%% have only medium complexity, and specifically loops are rare in the Diem
%% framework (loops were not discussed in this paper for reasons of space; we do
%% incorporate loop invariants for their verification). We anticipate the problem
%% of timeouts to hit us again down the road as the number of applications of \MVP
%% grow, and further refinement of the translation will be needed. Furthermore,
%% while currently there are no occurrences of false positives, we expect this problem
%% to hit as well.

%% The biggest obstacle for wider adoption, however, is seen in the complexity of
%% authoring and reading specifications, related to a phenomena we refer to as
%% ``specification boilerplate''. Pre/post condition specifications are often
%% verbose and difficult to read and write. At the time of this writing, function
%% specifications in the Diem framework are often of similar size than function
%% implementations.  A common problem is also sharing between aspects of specifications:
%% for example, if function |f| calls function |g| and the later function aborts, the
%% aborts condition propagates from |g| to the caller |f|. While the Move specification
%% language has mechanisms to express this kind of sharing by referring to the aborts
%% condition of |g| in the specification of |f|, this mechanism also makes specifications
%% less readable, and leaks implementation details into the specification of |f|.
%% We are currently experimenting with multiple strategies to cope with this, ranging
%% from a better methodology and language support to structure specifications, to
%% automatically inferring specs for some functions.



%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
