\UseRawInputEncoding
\documentclass[runningheads]{llncs}
%\sloppy

\input{prelude.tex}

%\overfullrule=1mm


\begin{document}


\author{
  David Dill\inst{1} \and Wolfgang Grieskamp\inst{1} \and \\ Junkil
  Park\inst{1} \and Shaz Qadeer\inst{1} \and Meng Xu\inst{1}
  \and Emma Zhong\inst{1}
}

\institute{Novi Research, Facebook}
\authorrunning{D. Dill, W. Grieskamp et. al.}

\title{Fast and Reliable Formal Verification of Smart Contracts with the Move Prover}

\maketitle
\begin{abstract}
  The Move Prover (\MVP) is a formal verifier for smart contracts
  written in the Move programming language. \MVP has an expressive
  specification language, and is fast and reliable enough that it can
  be run routinely by developers and in integration testing in a few
  minutes. Besides the simplicity of smart contracts and the Move
  language, three transformations are responsible for the practicality
  of \MVP: (1) an alias-free memory model, (2) fine-grained invariant
  checking, and (3) monomorphization.  The entirety of the Move code for
  the Diem blockchain has been extensively specified and can be
  completely verified by \MVP in a few minutes. Changes in the Diem
  framework must be successfully verified before being integrated into
  the open source repository on GitHub.
  \keywords{Smart contracts \and
    formal verification \and Move language \and Diem blockchain}
\end{abstract}


\input{intro.tex}
\input{move.tex}
\input{design.tex}
\input{analysis.tex}
\input{conclusion.tex}

%\vspace{-0.5ex}
\Paragraph{Acknowledgements}

This work would not have been possible without the many contributions of the
Move platform team and collaborators.  We specifically like to thank Bob Wilson,
Clark Barrett, Dario Russi, Jack Moffitt, Mathieu Baudet, Runtian Zhou, Sam
Blackshear, Tim Zakian, Todd Nowacki, Victor Gao, and Kevin Cheang.

\newpage
\bibliographystyle{splncs04}
\bibliography{biblio}

\appendix
\newpage
\input{function-injection-apx.tex}
\newpage
\input{corrected_example-apx.tex}


\end{document}


%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
