
\Section{Move and the Prover}

Move was developed for the Diem blockchain \cite{DIEM}, but its design is not
specific to blockchains.  A Move execution consists of a sequence of updates
evolving a \emph{global persistent memory state}, which we just call the
\emph{(global) memory}.
%% Updates are executed in a \emph{transactional} style.
Similar to other blockchains, updates are a series of atomic transactions.
All runtime errors result in a transaction abort, which does not change
the blockchain state except to transfer some currency (``gas'') from the account that
sent the transaction to pay for cost of executing the transaction.

The global memory is organized as a collection of resources, described by Move
structures (data types). A resource in memory is indexed by a pair of a type
(fully instantiated) and an address (for example the address of a user
account). For instance, the expression |exists<Coin<USD>>(addr)| will be true if
there is a value of type |Coin<USD>| stored at |addr|. As seen in this example,
Move uses type generics, and working with generic functions and types is rather
idiomatic for Move.  Notice that account addresses are not just arbitrary values
but have a specific role in Move's programming methodology related to access
control via the builtin type of \emph{signers}, as will be discussed later.

A Move application consists of a set of \emph{transaction scripts}. Each
script defines a Move function with input parameters but no output
parameters.  This function updates the global memory and may emit
event records. The execution of this function can abort because of an abort
instruction or implicitly because of a runtime error such as an out-of-bounds
vector index.

%% mechanism, in which case the memory stays unmodified. An environment emits a
%% sequence of calls to such scripts, thereby evolving the memory.

% A transaction script is written in Move as an imperative function which can read
% and write the global memory. Move uses a specific style of
% imperative programming based on \emph{borrow semantics} \cite{BORROW_SEM}, as
% popularized in the programming language Rust \cite{RUST,RUST_TYPES}. For the
% verification problem borrow semantics is very important.  While allowing
% references into structured data, those are guaranteed to be safe by the
% \emph{borrow checker} \cite{BORROW_CHECKER}, which is run during bytecode
% loading time, and which verification can assume. Furthermore, the notorious hard
% verification problem of aliasing of references in the presence of mutation is
% eliminated.  Mutation always starts from a root location either in global memory
% or on the execution stack, and while a tree starting from this root is mutated,
% no other access can happen anywhere in the tree. Intuitively, borrow semantics
% allows to move a mutation 'cursor' down the tree, which follows linear typing
% discipline. Because of this property, mutable reference parameters to functions
% can be converted to input/output parameters, and verification of Move can avoid
% the traditionally hard problems caused by aliasing of mutable references.


\Paragraph{Programming in Move}

\begin{Figure}
\caption{\label{fig:AccountDef} Account Example Program}
\begin{MoveBox}
module Account {
  struct Account has key {
    balance: u64,
  }

  fun withdraw(account: address, amount: u64) acquires Account {
    let balance = &mut borrow_global_mut<Account>(account).balance;
    assert(*balance >= amount, Errors::limit_exceeded());
    *balance = *balance - amount;
  }

  fun deposit(account: address, amount: u64) acquires Account {
    let balance = &mut borrow_global_mut<Account>(account).balance;
    assert(*balance <= Limits::max_u64() - amount, Errors::limit_exceeded());
    *balance = *balance + amount;
  }

  public(script) fun transfer(from: &signer, to: address, amount: u64)
  acquires Account {
    assert(Signer::address_of(from) != to, Errors::invalid_argument());
    withdraw(Signer::address_of(from), amount);
    deposit(to, amount);
  }
}
\end{MoveBox}
\end{Figure}

\noindent In Move, one defines transactions via so-called \emph{script functions} which
take a set of parameters.  Those functions can call other functions. Script and
regular functions are encapsulated in \emph{modules}. Move modules are also the
place where structs are defined. An illustration of a Move contract is given in
Fig.~\ref{fig:AccountDef} (for a more complete description see the Move
Book~\cite{MOVE_LANG_DEF}). The example is a simple account which holds a
balance, defined in the script function |transfer|.
%% Scripts get passed in so
Scripts generally have
\emph{signer} arguments, which are tokens which represent an account
address that has been authenticated by a cryptographic signature.
%% The caller of the script -- an external program -- has ensured that the
%% owner of the signer account address has agreed to execute this script.
%% Notice that in the code,
The |assert| statement in the example causes a Move transaction to abort
execution if the condition is not met.  Abortion can also happen implicitly; for
example, the expression |borrow_global_mut<T>(addr)| will abort if no resource
|T| exists at |addr|.

\Paragraph{Specifying in Move}

\begin{Figure}
\caption{\label{fig:AccountSpec} Account Example Specification}
\begin{MoveBox}
module Account {
  spec transfer {
    let from_addr = Signer::address_of(from);
    aborts_if from_addr == to;
    aborts_if bal(from_addr) < amount;
    aborts_if bal(to) + amount > Limits::max_u64();
    ensures bal(from_addr) == old(bal(from_addr)) - amount;
    ensures bal(to) == old(bal(to)) + amount;
  }

  spec fun bal(acc: address): u64 {
    global<Account>(acc).balance
  }

  invariant forall acc: address where exists<Account>(acc):
    bal(acc) >= AccountLimits::min_balance();

  invariant update forall acc: address where exists<Account>(acc):
    old(bal(acc)) - bal(acc) <= AccountLimits::max_decrease();
}
\end{MoveBox}
\end{Figure}

\noindent The specification language supports {\em Design By Contract}
\cite{DESIGN_BY_CONTRACT}. Developers can provide pre and post conditions for
functions, which include conditions over
%% (mutable)
parameters and global
memory. Developers can also provide invariants over data structures, as well as
the
%% (state-dependent)
contents of the global memory.
%
Universal and existential
quantification %% both
over bounded domains, such as like the indices of a vector, as well as
effectively unbounded domains, such as memory addresses and integers, are
supported.
%
Quantifiers make the verification problem undecidable
and cause difficulties with timeouts.
%
However, in practice, we notice that quantifiers have the advantage
of allowing more direct formalization of many properties, which increases the
clarity of specifications.
%% The latter makes the specification language very expressive, but also
%% renders the verification problem in theory undecidable (and in practice
%% dependent on heuristic decision procedures).

Fig.~\ref{fig:AccountSpec} illustrates the specification language by extending
the account example in Fig.~\ref{fig:AccountDef} (for the definition of the
specification language see \cite{MOVE_SPEC_LANG_DEF}). This adds the
specification of the |transfer| function, a helper function |bal| for use in
specs, and two global memory invariants. The first invariant states that a
balance can never drop underneath a certain minimum. The second invariant refers
to an update of global memory with pre and post state: the balance on an account
can never decrease in one step more than a certain amount.  Note that while the
Move programming language has only unsigned integers, the specification language
uses arbitrary precision signed integers, making it convenient to specify
something like |x + y <= limit|,
%% without need to worry about overflow.
without the complication of arithmetic overflow.

Specifications for the |withdraw| and |deposit| functions have been omitted in
this example.

\TODO{Dave}{Maybe we don't need to mention inlining? It's not
  necessary to understand the rest of the paper, and it doesn't fit in
  any obvious paragraph here.}

\MVP supports omitting specs for non-recursive functions, in
which case they are treated as being inlined at caller site.

%% A discerning reader may have noted that the program in Fig.~\ref{fig:AccountDef}
%% does not actually satisfy the specification in Fig.~\ref{fig:AccountSpec}. This
%% will be discussed in the next section.

% The constructs we have seen so far are only a subset of the available features
% of the Move specification language. Notably, the language supports the following
% additional features:

% \begin{itemize}
% \item Function preconditions via the |requires|-clause.
% \item Data invariants for |struct| types, as a predicate over the field values.
% \item Means to abstract commonly used specification fragments in so-called
%   \emph{specification schemas} which can then be included in other specification
%   blocks.
% \end{itemize}

\Paragraph{Running the Prover}
\label{sec:RunningProver}

\TODO{Dave}{If we need to cut for space, this can go. The error
  reporting doesn't support our main topic, which is how we made it
  fast and reliable.}

\MVP is fully automatic, like a type checker or linter, and is
expected to finish in a reasonable time, so it can be integrated in
the regular development workflow. Running \MVP on the |module Account| produces
multiple errors. The first is this one:

\begin{MoveDiag}
error: abort not covered by any of the `aborts_if` clauses
   -- account.move:24:3
   |
13 |       let balance = &mut borrow_global_mut<Account>(account).balance;
   |                          ----------------- abort happened here
   |
   =     at account.move:18: transfer
   =         from = signer{0x18be}
   =         to = 0x18bf
   =         amount = 147u8
   =     at ...
\end{MoveDiag}

\noindent \MVP detected that an implicit abort condition is missing in the
specification of the |withdraw| function. It prints the context of the error, as
well as an \emph{execution trace} which leads to the error. Values of variable
assignments from the counterexample found by the SMT solver are printed together
with the execution trace. Logically, the counterexample presents an
assignment to variables where the program fails to meet the specification. In
general, \MVP attempts to produce readable diagnostics for Move developers
without the need of understanding any internals of the prover.

The next errors produced are about the memory invariants in
Fig.~\ref{fig:AccountSpec}. Neither holds, but
%% Both of them do not hold;
we show only the first failure:

\begin{MoveDiag}
error: global memory invariant does not hold
   -- account.move:39:3
   |
39 |    invariant forall acc: address where exists<Account>(acc):
40 |      bal(acc) >= AccountLimits::min_balance();
   |
   =     at account.move:6: withdraw
   =         account = 0x0
   =         amount = 11u8
   =     at account.move:7: withdraw
   =         balance = &15u8
   =     at ...
\end{MoveDiag}

\noindent This happens because in the program in Fig.~\ref{fig:AccountDef}, we
%% did not made any
made no attempt to respect the limits in |min_balance()| and
|max_decrease()|.
The problem can be fixed by adding more |assert| statements to check that
the limits are met (see App.~\ref{sec:CorrectedAccount}).

%% We leave it open here (but covered it in
%% App.~\ref{sec:CorrectedAccount}) how to fix this problem, which requires
%% to add some more |assert| statements to the code and abort if the limits are not
%% met.

%% In reality,
The programs and specifications \MVP deals with are much larger than
this example. The conditions under which a transaction
in the Diem framework can abort typically involve dozens of individual predicates,
stemming from other functions called by this transaction. Moreover, there are
hundreds of memory invariants specified, encoding access control and other
requirements for the Diem blockchain.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
